---
title: "Class Reference"
author: "Marianeli De Leon"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this my personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  )
```
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```
```{r}
library(tidyverse)
```
 arrange(max_loan_amount)=acsending order
 command enter to read one line
 
 # mutate(new column=value)
 
 # "Maryland" is not "MARYLAND"
 
 in_out in or out of somewhere
 
 case_when= 
 
 filter=limits data to just one thing