---
title: "lab_08"
author: "marianeli de leon"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
library(tidyverse)
library(tidycensus)
census_api_key("23d86a0f3402e889b16419069579742ff2bf5476")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
ac5 <- load_variables(2019, "acs5", cache = TRUE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file
data<- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first in your list of ACS variables.
**A1**Baltimore City, MD had the lowest median age. 

```{r}
md <- get_acs(geography = "county",
              variables = c(medinage = "B01002_001"),
              state = "MD",
              year = 2019)


md <- md %>% arrange(estimate)
```

**Q2** Which Maryland counties have a median age of at least 50? Write code to isolate just those jurisdictions.
**A2**Worcester and Talbot Counties have a median age of at least 50. 

```{r}
md %>%
  group_by(estimate)%>%
  arrange(desc(estimate))
  
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? 

You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home 

and a second dataframe with the total population of 5-17 year-olds by county. 

For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable,

remove the `variable` columns in each using select and 

then join them based on GEOID and NAME. 

Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3**Prince George's, Montgomery, and Caroline counties have the highest percentage of 5-17 year olds who speak Spanish at home.

```{r}
spanish_5_17 <- get_acs(geography = "county",
              variables = c(medspanish = "B16007_004"),
              state = "MD",
              year = 2019)

total_5_17<- get_acs(geography = "county",
              variables = c(medtotal= "B16007_002"),
              state = "MD",
              year = 2019)

spanish_5_17 %>%
  inner_join(total_5_17, by=c("GEOID", "NAME")) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y)%>% 
  rename(spanish=estimate.x, total=estimate.y) %>%
  mutate(percent = (spanish/total)*100) %>% 
  arrange(desc(percent))
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4**Caroline County being one of the top percentages was interesting as I didn't think it would be anywhere near the top 5.  The counties are mainly located within Central Maryland and often have a large Hispanic population. Garrett county's numbers are extremely lower than the top 5 counties in Maryland and not known for a large Hispanic population. 

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? 

To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code.

In each dataframe, rename the `estimate` column and remove the `variable` column. 

You'll need to combine the two dataframes in order to calculate the percentage,

and then you can combine it with the PPP data. 

Remember: the PPP data may not have only 5-digit zip codes!
**A5** 72005 had the most removed PPP applications

```{r}
black_pop <- get_acs(geography = "zcta",
              variables = c(medsbla = "B01001B_001"),
              year = 2019)

total_zip<- get_acs(geography = "zcta",
              variables = c(medzip= "B01001_001"),
              year = 2019)

pop_data<-black_pop %>%
  inner_join(total_zip, by=c("GEOID", "NAME")) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y)%>% 
  rename(black=estimate.x, total=estimate.y) %>%
  mutate(percent = (black/total)*100) %>% 
  arrange(desc(percent))


data_clear<-data%>%
    mutate(zip5 = str_sub(zip, 1, 5))

top_zips <- data_clear %>%
 group_by(zip5) %>%
 summarise(count= n()) %>%
 arrange(desc(count))

final<-pop_data
inner_join(total_zip, by=c("GEOID", "NAME")) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y)%>% 
  rename(black=estimate.x, total_pop=estimate.y) %>%
  mutate(percent = (black/total)*100) %>% 
  arrange(desc(percent))

ppp_data <- total_zip%>%
  ppp_data<-get_acs(geography="zcta", variables = "NAME")%>%
zips_combo <- final %>% left_join(ppp_data, by=c('percent'='estimate'))

```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed applications in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6**Most of the top 10 zip codes had a majority of Black population and many of the zip codes are within the Southern U.S. areas that are on the poorer side and many of the zip codes don't seem too different. 

**Q7** What other columns from the PPP data could be useful in reporting on what happened to these removed loans?
**A7**The name and estimate columns could help in reporting.
