---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](https://www.propublica.org/article/over-a-dozen-black-and-latino-men-accused-a-cop-of-humiliating-invasive-strip-searches-the-nypd-kept-promoting-him)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
read_csv("data/CCRB-Complaint-Data_202007271729/allegations_202007271729.csv")
allegation_data<-read_csv("data/CCRB-Complaint-Data_202007271729/allegations_202007271729.csv")
cleaned_allegation_data<- allegation_data%>%
  distinct()
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at least one misconduct allegation that was substantiated by the CCRB, meaning that investigators amassed enough evidence of offenses, ranging from bad language to pistol whippings, to say that they happened and broke patrol guidelines. The most common involved improper property entries and searches.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
allegation_data%>%
  filter(str_detect(board_disposition,"^Subs"))%>%
  filter(rank_now=="Captain")%>%
  filter(complainant_gender=="Male")%>%
  filter(complainant_ethnicity=="Black", "Hispanic")%>%
  group_by(unique_mos_id, rank_abbrev_incident, rank_abbrev_now, complainant_ethnicity,complainant_gender)%>%
  summarise(
    count= n()
  )%>%
   arrange(desc(count))
  filter(rank_now == "Chiefs and other ranks" | rank_now == "Deputy Inspector" | rank_now == "Inspector") %>% 
  filter(complainant_gender=="Male")%>%
  filter(complainant_ethnicity=="Black" & "Hispanic")%>%
  select(unique_mos_id) %>% 
  distinct() %>% 
  group_by(unique_mos_id, rank_abbrev_incident, rank_abbrev_now, complainant_ethnicity,complainant_gender) %>% 
  summarize(count = n()) %>% 
  summarize(total = sum(count))
  
   filter(rank_now == "Chiefs and other ranks" | rank_now == "Deputy Inspector" | rank_now == "Inspector" & str_detect(board_disposition, "^Subs")) %>% 
 filter(complainant_gender=="Male")%>%
  filter(complainant_ethnicity=="Black" & "Hispanic")%>%
  group_by(allegation) %>% 
  summarize(count= n()) %>% 
  arrange(desc(count))

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Of at least 77 allegations made against him in 26 separate CCRB complaints, 29 were unsubstantiated; five were “unfounded,” meaning investigators concluded the incident never took place; and 27 were “exonerated,” meaning the conduct fell within patrol guidelines. Investigators have never proved that McCormack strip-searched men in the street, let alone penetrated them with his finger, though the city settled four lawsuits involving strip-search allegations against him for a total $184,500]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [An overwhelming majority of the city’s 36,000 officers do not have a single substantiated complaint on their records; only 1 in 9 does]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Data shows that while more than half of officers in the lowest rank are people of color, the higher ranks — above captain — are more than 75% white. A dozen of those high-ranking officers have three or more complaints with substantiated allegations; 10 are white. Five have reached the rank of deputy inspector, the lowest promoted at the commissioner’s discretion; all are white.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [McCormack has been accused of misconduct in at least 16 lawsuits, 11 of which settled for a combined $415,000]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-